Nextstrain: real-time tracking of virus evolution

### 2017-10-10 whitepaper.tex
Minor addition

\subsection*{A model for public sharing of data}
Nextstrain presents a single overview of both endemic viral disease (seasonal influenza, dengue) as well as emergent viral outbreaks (avian influenza, Zika, Ebola), all based upon the same underlying bioinformatics architecture.
As such, it is well positioned to respond to both the next outbreak and continued surveillance of current ones.
Analysis of such outbreaks relies on public sharing of data, and Nextstrain has the ability to automatically update sequences from a range of public databases and repositories.
Scientists are justifiably hesitant to cede control of their data, and we try to address these concerns by preventing access to the raw genome sequences, and by clearly indicating the source of each sequence.
We do, however provide the ability to download phylogenetic trees, metadata and screenshots, and private metadata can be added by users by simply dragging a CSV file onto the browser.
We believe this strikes a compromise between keeping certain data private and allowing the dissemination of results important to the wider scientific community, thereby encouraging collaboration between scientists.
Genomic epidemiology has the potential to inform the public, health organisations and scientists alike, a potential which is realised by sharing of data in real-time rather than retrospectively \citep{croucher2015application}.

>>>>

\subsection*{A model for public sharing of data}
Nextstrain presents a single overview of both endemic viral disease (seasonal influenza, dengue) as well as emergent viral outbreaks (avian influenza, Zika, Ebola), all based upon the same underlying bioinformatics architecture.
This same architecture should be fully extendable to bacterial systems in the future.
It is well positioned to respond to both the next outbreak and continued surveillance of current ones.
Analysis of such outbreaks relies on public sharing of data, and Nextstrain has the ability to automatically update sequences from a range of public databases and repositories.
Scientists are justifiably hesitant to cede control of their data, and we try to address these concerns by preventing access to the raw genome sequences, and by clearly indicating the source of each sequence.
We do, however provide the ability to download phylogenetic trees, metadata and screenshots, and private metadata can be added by users by simply dragging a CSV file onto the browser.
We believe this strikes a compromise between keeping certain data private and allowing the dissemination of results important to the wider scientific community, thereby encouraging collaboration between scientists.
Genomic epidemiology has the potential to inform the public, health organisations and scientists alike, a potential realised by sharing of data in real-time rather than retrospectively \citep{croucher2015application}.
